---
title: "Analyzing Q15 by Q3"
author: "Jarad Niemi"
date: "November 10, 2016"
output: html_document
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Analyzing Q15 by Q3}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the data and remove all questions except Q3 and Q15 and remove Q15p as it
is an "Other" category. 

```{r load_data, message=FALSE}
library(dplyr)

d <- STRIPSArbuckle::survey_responses %>% 
  dplyr::select(CaseID,
         starts_with("q3"), 
         starts_with("q15"),
         -starts_with("q15p")) %>%
  tidyr::gather(question, response, -CaseID, -q3) %>%
  rename(farmer = q3)
```

Compute sufficient statistics for the data, i.e. the number of respondents for 
each category (No Response, No Priority, Slight Priority, Moderate Priority, 
High Priority, and Very High Priority) for all parts of question 15 stratified
by their answer to question 3. 

```{r summarize_data}
ordered_responses = c("No response",
                      "No Priority",
                      "Slight Priority",
                      "Moderate Priority",
                      "High Priority",
                      "Very High Priority")

s = d %>%
  filter(!is.na(farmer)) %>%
  mutate(question = substr(question, 1,4),
         response = ifelse(is.na(response), "No response", response)) %>%
  group_by(farmer, question, response) %>%
  summarize(n = length(CaseID)) %>%
  ungroup() %>%
  mutate(response = factor(response, levels=ordered_responses))
```

For each question, run a t-test by assuming numeric values for the responses
that are consecutive integers (0 to 4 below). 

```{r t_test}
# Run t test to get t statistic and p value
make_data <- function(n) {
  data.frame(y=rep(0:4,n)) # Converts 
}

my_t.test <- function(d) {
  test = t.test(make_data(d$No),
                make_data(d$Yes),
                var.equal = TRUE)
  data.frame(t = formatC(test$statistic, 2, format='f'),
             p = formatC(test$p.value, 3, format='f'))
}

test = s %>%
  tidyr::spread(farmer,n) %>%
  arrange(question, response) %>%
  filter(response != "No response") %>%
  group_by(question) %>%
  do(my_t.test(.))
```

```{r}
# Create data.frame with nice question formatting
pretty_names = 
  c("Increase recreation opportunities",
    "Beautification",
    "Increase crop production",
    "Increase livestock production",
    "Improve flood control",
    "Drinking water quality",
    "Swimming water quality",
    "Aquatic life water quality",
    "Restore native prairie",
    "Restore wetlands",
    "Reduce greenhouse gases",
    "Increase rural job opportunities",
    "Increase tourism",
    "Improve game habitat",
    "Improve non-game habitat")

names(pretty_names) <- paste0("q15",letters[1:15])
```


Combine the counts with the t statistic and p value to provide a parsimonious 
summary of the results. 

```{r print_results, results='asis'}
ordered_colnames = paste(rep(c("No","Yes"), each=length(ordered_responses)), 
                         ordered_responses,
                         sep="_")

tmp_for_table = s %>%
  mutate(column_name = factor(paste(farmer,response,sep="_"), 
                              levels = ordered_colnames)) %>%
  select(question, column_name, n) 

for_table <- tmp_for_table %>%
  tidyr::spread(column_name, n) %>%
  left_join(test, by="question") %>%
  mutate(question = plyr::revalue(question, pretty_names))

names(for_table) <- gsub("No_", "", names(for_table))
names(for_table) <- gsub("Yes_","", names(for_table))


htmlTable::htmlTable(for_table,
                     rnames = FALSE,
                     cgroup = c("","Non-farmer","Farmer","",""),
                     n.cgroup = c(1,6,6,1,1),
                     align = "l|rrrrrr|rrrrrr|rr")
```